---
title: "Pig Cell Atlas - MafB"
author: "Domien"
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  header-includes: "-| ```{=latex} \\usepackage{fvextra} \\DefineVerbatimEnvironment{Highlighting}{Verbatim}{
    breaksymbolleft={}, showspaces = false, showtabs = false, breaklines, commandchars=\\\\\\{\\}
    } ```"
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

\newpage

# Introduction

To investigate the conservedness of MafB to imprint Mac identity across species we quantified the correlation between Mafb expression and the expression of macrophage sigature genes and MafB target genes in single cell atlases of mouse (Tabula Muris Senis), human (Tabula Sapiens), lemur (Tabula Microcebus), pig (Pig Cell Atlas), African clawed frog (Xenopus Cell Landscape) and zebrafish (Zebrafish Cell Landscape, ZCL).

# Load packages

```{r}
suppressMessages({
library(SeuratObject)
library(Seurat)
library(reticulate)
library(sceasy)
library(ggplot2)
library(ggrastr)
library(readxl)
})
```

# Download Pig Cell Atlas Seurat object from Pigatlas server
```{bash, eval = FALSE}
https://dreamapp.biomed.au.dk/pigatlas/
```

# Load Pig Cell Atlas Seurat object

```{r}
PCA <- readRDS("pig_atlas_20221014.rds")
```

# Visualize clusters

```{r fig.height=4, fig.width=4}
p1 <- DimPlot(PCA, reduction = "tsne", raster = F)+
  theme(legend.position='none')+ggtitle("Pig Cell Atlas")

rasterize(p1, layers='Point', dpi=1200)
```

# Calculate correlation between MAFB expression and Mac signature score

```{r fig.height=4, fig.width=4.5}
FeaturePlot(PCA, features = "MAFB", reduction = "tsne", raster = T)
```

```{r}
Mac_sign <- read_excel("Mac_sign_pig.xlsx")

Mac_sign <- Mac_sign$ortholog_name

Mac_sign <- intersect(Mac_sign, rownames(PCA))

PCA <- AddModuleScore(PCA, features = list(c(Mac_sign)), name = "Mac_sign")
```

```{r fig.height=4, fig.width=4}
FeatureScatter(PCA, feature1 = "MAFB", feature2 = "Mac_sign1", raster = T)+
  geom_smooth(method='lm', colour = "black")+
  theme(legend.position='none')
```

```{r}
corr <- FetchData(PCA, vars = "MAFB") 
corr <- cbind(corr, FetchData(PCA, vars = "Mac_sign1"))

cor.test(corr$MAFB, corr$Mac_sign1, method = "pearson")
```

```{r fig.height=4, fig.width=4}
p2 <- ggplot(corr, aes(x = MAFB, y = Mac_sign1))+
  geom_point(size = 1, colour="grey")+
  geom_smooth(method='lm', colour = "black")+
  xlab("MAFB Expression")+
  ylab("Macrophage signature score")+
  theme_classic()
rasterize(p2, layers='Point', dpi=600)
```

# Calculate correlation between MAFB expression and MafB target gene signature score

```{r}
MafB_target_sign <- read_excel("MafB_target_pig.xlsx")


MafB_target_sign <- MafB_target_sign$ortholog_name

MafB_target_sign <- unique(MafB_target_sign)

MafB_target_sign <- intersect(MafB_target_sign, rownames(PCA))

PCA <- AddModuleScore(PCA, features = list(c(MafB_target_sign)), name = "MafB_target_sign")
```

```{r fig.height=4, fig.width=4.5}
FeatureScatter(PCA, feature1 = "MAFB", feature2 = "MafB_target_sign1", raster = F)+
  geom_smooth(method='lm', colour = "black")+
  theme(legend.position='none')
```

```{r}
corr <- cbind(corr, FetchData(PCA, vars = "MafB_target_sign1"))

cor.test(corr$MAFB, corr$MafB_target_sign1, method = "pearson")
```

```{r fig.height=4, fig.width=4}
p3 <- ggplot(corr, aes(x = MAFB, y = MafB_target_sign1))+
  geom_point(size = 1, colour="grey")+
  geom_smooth(method='lm', colour = "black")+
  xlab("MAFB Expression")+
  ylab("MafB target gene signature score")+
  theme_classic()
rasterize(p3, layers='Point', dpi=600)
```

```{r}
sessionInfo()
```




