---
title: "Tabula Muris Senis Myeloid"
author: "Domien Vanneste"
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
  header-includes:
    -|
    ```{=latex}
    \usepackage{fvextra}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      breaksymbolleft={}, 
      showspaces = false,
      showtabs = false,
      breaklines,
      commandchars=\\\{\}
    }
    ```
---


```{r include=FALSE}
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

# Load packages

```{r}
suppressMessages({
library(Seurat)
library(ggplot2)
library(reticulate)
library(sceasy)
library(ggsci)
library("scales")
library(ggrastr)
library(OmnipathR)
library(decoupleR)
library(dplyr)
library(tibble)
library(tidyr)
library(patchwork)
})
```

# Download Tabula Muris Senis scanpy object from AWS server
```{bash, eval = FALSE}
~$ aws s3 cp --no-sign-request s3://czb-tabula-muris-senis/Data-objects/tabula-muris-senis-bbknn-processed-official-annotations.h5ad /home/uliege/Documents/BioInformatics/TabulaMurisSenis
download: s3://czb-tabula-muris-senis/Data-objects/tabula-muris-senis-bbknn-processed-official-annotations.h5ad to Documents/BioInformatics/TabulaMurisSenis/tabula-muris-senis-bbknn-processed-official-annotations.h5ad
```


# Convert Scanpy object to Seurat object
```{r, eval = FALSE}
sceasy::convertFormat("tabula-muris-senis-bbknn-processed-official-annotations.h5ad", from="anndata", to="seurat",
                       outFile='tbms.rds')
```

# Load Seurat object
```{r}
tbms <- readRDS(file = "tbms.rds")
```


```{r}
DimPlot(tbms, group.by = "louvain", label = T, raster = T)
```

# Subset myeloid cells

Myeloid cell clusters were identified based on the combined expression of Ptprc (CD45) anf Lyz2 (LysM).

```{r}
FeaturePlot(tbms, features = "Ptprc", raster = T)
```

```{r}
VlnPlot(tbms, features = "Ptprc", group.by = "louvain", pt.size = 0)
```

```{r}
FeaturePlot(tbms, features = "Lyz2", raster = T)
```

```{r}
VlnPlot(tbms, features = "Lyz2", group.by = "louvain", pt.size = 0)
```

```{r}
Myeloid_Cells <-c(2, 3, 9, 11, 12, 15, 16, 36, 43, 45, 48, 52)

tbms_myeloid <- subset(tbms, subset = louvain %in% Myeloid_Cells)
```

```{r}
DimPlot(tbms_myeloid, raster = T)+theme(legend.position='none')
```

# Normalization and scaling of myeloid cells

```{r}
tbms_myeloid <- FindVariableFeatures(tbms_myeloid, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(tbms_myeloid)
tbms_myeloid <- ScaleData(tbms_myeloid, features = all.genes)
tbms_myeloid <- RunPCA(tbms_myeloid, features = VariableFeatures(object = tbms_myeloid))
ElbowPlot(tbms_myeloid, ndims = 50)
```

# Clustering of myeloid cells
```{r}
tbms_myeloid <- FindNeighbors(tbms_myeloid, dims = 1:7)
tbms_myeloid <- FindClusters(tbms_myeloid, resolution = 0.35)
tbms_myeloid <- RunUMAP(tbms_myeloid, dims = 1:7)
DimPlot(tbms_myeloid, label = T, raster = T)
```


```{r}
DimPlot(tbms_myeloid, group.by = "tissue", label = T, raster = T)
```

# Removal of contaminating cells

```{r}
cluster16.markers <- FindMarkers(tbms_myeloid, ident.1 = 16, only.pos = TRUE)
```
Cluster 16 are proliferating cells

```{r}
cluster17.markers <- FindMarkers(tbms_myeloid, ident.1 = 17, only.pos = TRUE)
```
Cluster 17 are mast cells

```{r}
tbms_myeloid <- subset(tbms_myeloid, idents = c(16, 17), invert = T)
```


```{r}
DimPlot(tbms_myeloid, label = T, raster = T)
```

# Assigning celltypes to clusters

```{r}
tbms_myeloid$cell_type <- factor(Idents(tbms_myeloid), labels = c("cMo", "Mature Neu", "Immature Neu", "cDC", "MG", "preNeu", "RTM", "MG", "GMP", "cMoP", "KC", "pMo", "proNeu", "pDC", "AM", "IM"))

tbms_myeloid$cell_type <- factor(tbms_myeloid$cell_type, levels = c("cMo", "Mature Neu", "Immature Neu", "cDC", "MG", "preNeu", "RTM", "GMP", "cMoP", "KC", "pMo", "proNeu", "pDC", "AM", "IM"))

Idents(tbms_myeloid) <- "cell_type"

DimPlot(tbms_myeloid, label = T, raster = T)
```

```{r}
tbms_myeloid$cell_type <- factor(tbms_myeloid$cell_type, levels = c("GMP", "proNeu", "preNeu", "Immature Neu", "Mature Neu", "cMoP", "cMo", "pMo", "cDC", "pDC", "AM", "IM", "KC", "RTM", "MG"))

Idents(tbms_myeloid) <- "cell_type"

DimPlot(tbms_myeloid, raster = T)
```


```{r}
show_col(pal_igv("default")(15))
```

```{r}
cols <- c(pal_igv("default")(15))

DimPlot(tbms_myeloid, label = T, cols = cols, group.by = "cell_type", raster = T)
```

```{r}
DimPlot(tbms_myeloid, cols = cols, group.by = "cell_type", raster = T)+ggtitle("Tabula Muris Senis Myeloid")
```


# Viualize Mafb expression

```{r}
FeaturePlot(tbms_myeloid, features = "Mafb", max.cutoff = 'q90', cols = c("lightgrey","red"), raster = T)+ggtitle("Mafb expression")
```


```{r}
VlnPlot(tbms_myeloid, features = "Mafb", pt.size = 0, cols = cols)+NoLegend()+ggtitle("Mafb expression")
```


# Calculate and viualize MafB activity with decoupleR

## Load CollecTRI network

CollecTRI is a comprehensive resource containing a curated collection of TFs and their transcriptional targets.

```{r}
net <- get_collectri(organism="mouse", split_complexes=FALSE)
net
```

## Activity inference with Univariate Linear Model (ULM)

To infer TF enrichment scores we will run the Univariate Linear Model (ulm) method. For each sample in our dataset (mat) and each TF in our network (net), it fits a linear model that predicts the observed gene expression based solely on the TF’s TF-Gene interaction weights. Once fitted, the obtained t-value of the slope is the score. If it is positive, we interpret that the TF is active and if it is negative we interpret that it is inactive.

```{r}
# Extract the normalized log-transformed counts
mat <- as.matrix(tbms_myeloid@assays$RNA@data)

# Run ulm
acts <- run_ulm(mat=mat, net=net, .source='source', .target='target', .mor='mor', minsize = 5)

acts
```

## Visualize

```{r}
# Extract ulm and store it in tfsulm in seurat object
tbms_myeloid[['tfsulm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = tbms_myeloid) <- "tfsulm"

# Scale the data
tbms_myeloid <- ScaleData(tbms_myeloid)
tbms_myeloid@assays$tfsulm@data <- tbms_myeloid@assays$tfsulm@scale.data
```

```{r}
FeaturePlot(tbms_myeloid, features = "Mafb", min.cutoff = 'q10', max.cutoff = 'q90', cols = c("lightgrey", "red"), raster = T)+ggtitle("MafB activity")
```


```{r}
VlnPlot(tbms_myeloid, features = "Mafb", pt.size = 0, cols = cols)+NoLegend()+ggtitle("MafB activity")
```


# Session information 

R sesssion: 

```{r}
sessionInfo()
```











